---
title: "Task 6 - Data analysis"
output: html_notebook
---

```{r}
# load in libraries
library(tidyverse)
library(here)

```

```{r}
# read in cleaned data set
dog_survey <- read_csv(here("clean_data/clean_data.csv"))

dog_survey

```

--------------------------------------------------------------------------------
## Analysis Questions


# The client only counts a valid email address as one ending in ‘.com’. How many survey results have a valid email address?

```{r}

dog_survey %>% 
  mutate(email_dot_com = str_detect(email, "[.][com]{3}$")) %>%
  filter(email_dot_com == T) %>% 
  distinct(id) %>% 
  summarise(dot_com_emails = n())

```
There are 181 email addresses ending with ".com".


# What’s the average amount spent on dog food for each dog size?

[Assumption] Dog sizes have been categorised into Small, Medium and Large in the
cleaning script. In the case of an owner having multiple dogs it has been
assumed that the cost of dog food is split evenly between each dog.

```{r}

dog_survey %>% 
  filter(!is.na(dog_size)) %>% 
  group_by(dog_size) %>% 
  summarise(spent_on_food = round(mean(amount_spent_on_dog_food, 
                                       na.rm = T), 2)) %>% 
  arrange(desc(spent_on_food))

```
On average £52.03 is spent on food for small dogs, the highest of the
categories, followed by £50.93 for large dogs and £47.26 for medium-sized dogs.


# For owners whose surname starts with a letter in the second half of the alphabet (N onwards) what is the average age of their dog?

```{r}
dog_survey %>% 
  filter(str_detect(last_name, "^[N-Z]")) %>% 
  summarise(avg_age_of_dog = round(mean(dog_age, na.rm = T),1))


```
The average age of dogs with owner's surname in the second half of the alphabet
is 52.9 years.

# The dog_age column is the age in dog years. If the conversion is 1 human year = 6 dog years, then what is the average human age for dogs of each gender?

```{r}
dog_survey %>% 
  mutate(human_years = dog_age / 6) %>% 
  group_by(dog_gender) %>% 
  summarise(avg_human_years = round(mean(human_years, na.rm = T), 1)) %>%
  filter(!is.na(dog_gender))

```
The average age of male dogs is 8.3 years and for females is 7.8 years

# Create a plot of results of question 4.

```{r}
dog_survey %>%
  mutate(human_years = dog_age / 6) %>% 
  group_by(dog_gender) %>% 
  summarise(avg_human_years = round(mean(human_years, na.rm = T), 2)) %>%
  filter(!is.na(dog_gender)) %>% 
  ggplot(aes(x = dog_gender, y = avg_human_years)) +
  geom_bar(stat="identity",
           width = 0.5) +
  labs(x = "Gender", y = "Average Dog Age (years)")

```
