---
title: "R Notebook"
output: html_notebook
---

```{r}

library(tidyverse)
library(janitor)
library(here)

```

```{r}

dog_survey <- read_csv(here("raw_data/dog_survey.csv")) %>% clean_names()

```

```{r}

dog_survey %>% 
  distinct()

```
### Cleaning script

```{r}

dogs <- dog_survey %>%
  select(-(x10:x11)) %>%
  distinct() %>% 
  mutate(dog_gender = recode(dog_gender, "1 male and 1 female" = "m,f"),
         dog_age = recode(dog_age, "5 and 4" = "5,4")) %>% 
  mutate(dog_size = str_split(dog_size, ","),
         dog_gender = str_split(dog_gender, ","),
         dog_age = str_split(dog_age, ",")) %>%
  unnest(cols = c(dog_size, dog_gender, dog_age)) %>% 
  mutate(dog_gender = str_to_lower(dog_gender)) %>%
  mutate(dog_gender = case_when(
    dog_gender %in% c("male", "m")  ~ "M",
    dog_gender %in% c("female","f", "femlae")  ~ "F"
  )) %>% 
  mutate(dog_age = str_extract(dog_age, "[0-9]+"),
         dog_age = as.numeric(dog_age)) %>% 
  mutate(dog_size = case_when(
    dog_size %in% c("XS", "S", "Smallish")  ~ "S",
    dog_size %in% c("M", "Medium sized")  ~ "M",
    dog_size %in% c("XL", "L", "large")  ~ "L"
  )) %>% 
  mutate(amount_spent_on_dog_food = str_extract(amount_spent_on_dog_food,
                                                "£[0-9]+.[0-9]+"),# good enough
         amount_spent_on_dog_food = str_remove(amount_spent_on_dog_food, "£"),
         amount_spent_on_dog_food = as.numeric(amount_spent_on_dog_food)) 

```

The client only counts a valid email address as one ending in ‘.com’. How many
survey results have a valid email address?

```{r}

dogs %>% 
  mutate(email_dot_com = str_detect(email, "[.][com]{3}$")) %>%
  filter(email_dot_com == T) %>% 
  distinct(id) %>% 
  summarise(dot_com_emails = n())


```

What’s the average amount spent on dog food for each dog size?

```{r}

dogs %>% 
  group_by(id) %>% 
  mutate(count_of_dogs = n()) %>% 
  filter(count_of_dogs == 1) %>% 
  ungroup() %>% 
  group_by(dog_size) %>% 
  summarise(spent_on_food = mean(amount_spent_on_dog_food, na.rm = T), n())

```



For owners whose surname starts with a letter in the second half of the alphabet
(N onwards) what is the average age of their dog?

```{r}
dogs %>% 
  filter(str_detect(last_name, "^[N-Z]")) %>% 
  summarise(avg_age_of_dog = mean(dog_age, na.rm = T))


```


The dog_age column is the age in dog years. If the conversion is 1 human year =
6 dog years, then what is the average human age for dogs of each gender?

```{r}
dogs %>% 
  mutate(human_years = dog_age / 6) %>% 
  group_by(dog_gender) %>% 
  summarise(avg_human_years = mean(human_years, na.rm = T)) %>%
  filter(!is.na(dog_gender))


```




Create a plot of results of question 4.