---
title: "Task 5 - Data analysis"
output: html_notebook
---

```{r}

library(tidyverse)
library(here)

```

# Read in cleaned data set
 
```{r}

rwa <- read_csv(here("clean_data/clean_data.csv"))

rwa

```

--------------------------------------------------------------------------------

Analysis questions:

# What’s the average RWA score for each gender?

```{r}
rwa %>% 
  filter(!is.na(gender)) %>% 
  group_by(gender) %>% 
  summarise(avg_rwa_score = round(mean(rwa_score), 2)) %>% 
  arrange(desc(avg_rwa_score))

```
Average RWA scores for males, females and other are 3.12, 2.47 and 1.89,
respectively.

# What’s the average RWA score for left handed people vs. right handed people?

```{r}

rwa %>% 
  filter(hand == "Right" | hand == "Left") %>% 
  group_by(hand) %>% 
  summarise(avg_rwa_score = round(mean(rwa_score), 3))

```
The average RWS score for left- and right-handed people are extremely similar, 
2.856 for left-handed people and 2.587 for right-handed people. Three decimal 
places are presented to show how similar the values are.


# What’s the average family size for each type of childhood?

[Assumption] "Type" of childhood has been interpreted as which type of area did
the person live in when they were a child, which relates to the column labelled
"Urban".

```{r}

rwa %>% 
  filter(!is.na(urban)) %>% 
  group_by(urban) %>% 
  summarise(avg_family_size = round(mean(familysize), 2)) %>% 
  arrange(desc(avg_family_size))

```
The data shows that family size tended to be larger in rural areas (2.79), 
followed by suburban areas (2.62) and lastly urban areas (2.5).


# What’s the average time to take the test for each education level?

[Assumption] "testelapse" values taken to be in units of seconds.

```{r}

rwa %>% 
  filter(!is.na(education_level)) %>% 
  group_by(education_level) %>% 
  summarise(avg_test_time = round(mean(testelapse), 0)) %>% 
  arrange(avg_test_time)

```
The results show that people with graduate degrees had the overall fastest time
to complete the test, on average. This was followed by people who dropped out of
high school, then high school graduates and lastly those with a university
degree.


# Create a plot of results of question 4.

```{r}

rwa %>% 
  filter(!is.na(education_level)) %>% 
  group_by(education, education_level) %>% 
  summarise(avg_test_time = round(mean(testelapse), 0)) %>% 
  ggplot(aes(x = education_level, y = avg_test_time)) + 
  geom_bar(stat = "identity") +
  labs(y= "Average Test Time (s)", x = "Education")

```
[don't quite know how to put the education levels in ascending order]

# What’s the average RWA score for people aged:
Under 18
18 to 25
26 to 40
41 to 60
Over 60

```{r}

rwa %>% 
  # categorise each numerical age value to an age range
  mutate(age_range = case_when(
    age < 18 ~ "Under 18",
    age <= 25 ~ "18 - 25",
    age <= 40 ~ "26 - 40",
    age <= 60 ~ "40 - 60",
    age > 60 ~ "Over 60"
  )) %>% 
  # calcualte the average RWA score for each age range
  group_by(age_range) %>% 
  summarise(avg_rwa_score = round(mean(rwa_score), 2)) %>% 
  arrange(desc(avg_rwa_score))

```
The output shows that the average RWA score decreases with age. Under 18s have
the highest score with 3.36 followed by the 18 - 25, 26 - 40, 40 - 60 and over
60s category with average RWA scores of 3.15, 2.61, 2.54 and 2.40, respectively.

