---
title: "Task 4 - Data analysis"
output: html_notebook
---

# Load in libraries

```{r}

library(tidyverse)
library(here)

```

# Read in cleaned dataset

```{r}

candy <- read_csv(here("clean_data/candy_cleaned"))

candy

```

--------------------------------------------------------------------------------

## Analysis questions

# What is the total number of candy ratings given across the three years. (Number of candy ratings, not the number of raters. Don’t count missing values)

```{r}

# since ratings with NA values is already removed via the cleaning script, the 
# number of ratings is simplified and can be calculated as follows:

candy %>%
  summarise(num_ratings = n())

```

# What was the average age of people who are going out trick or treating?

```{r}

candy %>% 
  filter(going_out == "Yes") %>%
  # group by id - there are multiple rows per id since data is in 'long' format
  group_by(id) %>%
  # avg age per id is just the age of the survey respondent
  summarise(avg_age = mean(age)) %>% 
  # average age across every survey respondent
  summarise(avg_age_pop = round(mean(avg_age, na.rm = T), 1))

```
The average age of the survey respondent who is going out trick or treating is
34.9 (to 1d.p.).

# What was the average age of people who are not going trick or treating?

```{r}

candy %>% 
  filter(going_out == "No") %>%
  group_by(id) %>%
  summarise(avg_age = mean(age)) %>% 
  summarise(avg_age_pop = round(mean(avg_age, na.rm = T), 1))

```
The average age of the survey respondent who is not going out trick or treating
is 39 (to 1d.p.). Note the decimal place has been dropped since value has been
rounded to a whole number.

# For each of joy, despair and meh, which candy bar received the most of these ratings?

```{r}

candy %>%
  group_by(rating, candy_name) %>% 
  summarise(num_ratings = n()) %>% 
  slice_max(num_ratings) 

```
Reese's Peanut Butter Cups received the highest number of JOY ratings (7,353),
Box o' raisins received the highest number of DESPAIR ratings (6,567) and
Lollipops received the highest number of MEH ratings (1,557)

# How many people rated Starburst as despair?

```{r}

candy %>% 
  filter(candy_name == "starburst",
         rating == "DESPAIR") %>% 
  summarise(count = n())
  
```
A total of 1,986 survey respondents rated Starbursts as "DESPAIR".

---For the next three questions, count despair as -1, joy as +1, and meh as 0.--

```{r}
# add column to dataset with scores for each rating
candy_scored <- candy %>% 
  mutate(score = case_when(
    rating == "JOY" ~ 1,
    rating == "DESPAIR" ~ -1,
    TRUE ~ 0
  ))

```

# What was the most popular candy bar by this rating system for each gender in the dataset?

```{r}

candy_scored %>%
  filter(!is.na(gender)) %>% 
  group_by(gender, candy_name) %>%
  summarise(total_score = sum(score)) %>% 
  slice_max(total_score)

```
All responses to gender were retained in the dataset as the question did not 
specify if genders other than Male/Female were to be removed.

Both males and females rated Reese's Peanut Butter Cups the highest. Kit Kat and
Twix were rated highest for people responding "I'd rather not say" or "Other".

# What was the most popular candy bar in each year?

```{r}

candy_scored %>% 
  group_by(year, candy_name) %>% 
  summarise(total_score = sum(score)) %>% 
  slice_max(total_score)

```
The most popular candy in 2015 and 2017 was Reese's Peanut Butter Cups, in 2016
the most popular candy was Kit Kat.

# What was the most popular candy bar by this rating for people in US, Canada, UK, and all other countries?

```{r}

candy_scored %>%
  filter(!is.na(country)) %>% 
  group_by(country, candy_name) %>% 
  # max total_score is the best, however if there is a tie then the candy with
  # the lowest number of votes wins (i.e. highest proportion of "JOY" votes)
  summarise(total_score = sum(score),
            num_votes = n()) %>%
  arrange(desc(total_score), num_votes) %>%
  slice_max(total_score, with_ties = F) %>% 
  select(country, candy_name)

```
The most popular candy in USA was Reese's Peanut Butter Cups, in the UK it was 
Rolos. Kit Kat was the most popular candy in both Canada and Other countries.