---
title: "R Notebook"
output: html_notebook
---

```{r}

library(tidyverse)
library(readxl)
library(janitor)
library(here)
library(assertr)


```

```{r}

candy_2015 <- read_xlsx("raw_data/boing-boing-candy-2015.xlsx") %>%
  clean_names()
candy_2016 <- read_xlsx("raw_data/boing-boing-candy-2016.xlsx") %>%
  clean_names()
candy_2017 <- read_xlsx("raw_data/boing-boing-candy-2017.xlsx") %>%
  clean_names()

```

```{r}

candy_2015
candy_2016
candy_2017


```




```{r}

names_2015 <- names(candy_2015)
names_2016 <- names(candy_2016)
names_2017 <- names(candy_2017)

```

### Clean 2015 data

```{r}

candy_cleaned_2015 <- candy_2015 %>% 
  rename(age = how_old_are_you, 
         going_out = are_you_going_actually_going_trick_or_treating_yourself) %>% 
  mutate(year = str_extract(timestamp, '[0-9]{1,4}'), .after = timestamp) %>%
  mutate(id = row_number(timestamp) + 1e6) %>% 
  mutate(country = NA_character_, .after = age) %>% 
  #mutate(is_going_out = ifelse(is_going_out == "Yes",T, F)) %>% 
  select(id, year:york_peppermint_patties, necco_wafers) %>% 
  # replace all non integer age inputs as NA, convert values to integers
  mutate(age = as.integer(age), year = as.integer(year)) %>% 
  pivot_longer(butterfinger:necco_wafers, names_to = "candy_name", values_to = "rating") %>% 
  select(id, year, going_out, age, country, candy_name, rating)


# maybe pivot_longer for combining three datasets
candy_cleaned_2015
```

### Clean 2016 data

```{r}

candy_cleaned_2016 <- candy_2016 %>% 
  rename(going_out = are_you_going_actually_going_trick_or_treating_yourself,
         age = how_old_are_you,
         country = which_country_do_you_live_in,
         gender = your_gender) %>%
  mutate(id = row_number(timestamp) + 2e6, .before = timestamp) %>% 
  mutate(year = str_extract(timestamp, '[0-9]{1,4}'), .after = timestamp) %>%
  clean_country_names() %>% 
  select(id, year:york_peppermint_patties, -gender, -which_state_province_county_do_you_live_in) %>% 
  # replace all non integer age inputs as NA, convert values to integers
  mutate(age = as.integer(age), year = as.integer(year)) %>% 
  #pivot_longer(x100_grand_bar:york_peppermint_patties, names_to = "candy_name", values_to = "rating")
  pivot_longer(x100_grand_bar:york_peppermint_patties, names_to = "candy_name", values_to = "rating") %>% 
  select(id, year, going_out, age, country, candy_name, rating)


candy_cleaned_2016

```

```{r}

candy_cleaned_2017 <- candy_2017 %>% 
  rename(id = internal_id) %>% 
 # rename(id = internal_id) %>% 
  pivot_longer(q1_going_out:q11_day, names_to = "col_names", values_to = "value") %>%
  select(id, col_names, value) %>%
  mutate(col_names = str_remove(col_names, "q[0-9]_")) %>%
  pivot_wider(names_from = col_names, values_from = value) %>%
  clean_names() %>%
  mutate(year = as.integer(2017), .after = id) %>%
  mutate(age = as.integer(age)) %>% 
  clean_country_names() %>% 
  pivot_longer(x100_grand_bar:york_peppermint_patties, names_to = "candy_name", values_to = "rating") %>% 
  select(id, year, going_out, age, country, candy_name, rating)

candy_cleaned_2017
```

Summary of countries
No cleaning = 118 unique countries
Remove non character values = 
Lowercase = 96 unique countries

```{r}

clean_country_names <- function(dataframe){
  
us_list <- c("merica", "ahemamerca", "alaska", "america","california","murica",
             "murrika","newjersey","newyork","northcarolina","pittsburgh",
             "trumpistan", "unhingedstates", "uniedstates","unitestates",
             "unitesstates", "uniteds", "us", "theunitedstates",
             "theunitedstatesofamerica", "unitedsates", "unitedstaes",
             "ipretendtobefromcanada,butiamreallyfromtheunitedstates",
             "unitedstate", "unitedstatea", "unitedstated", "usofa", "ussa",
             "ud", "USA", "sub-canadiannorthamericamerica",
             "theyooessofaaayyyyyy", "unitsstates")

uk_list <- c("endland", "england", "scotland", "uk", "unitedkingdom",
             "unitedkindom")

canada_list <- c("can", "canada")

exclude_list <- c("a", "canae", "cascadia", "earth", "fearandloathing",
                  "idontknowanymore", "insanitylately", "namerica", "narnia",
                  "sovietcanuckistan", "subscribetodmuzonyoutube", "denial",
                  "godscountry", "neverland", "oneofthebestones", "seeabove",
                  "somewhere", "eua", "thereisntoneforoldmen",
                  "therepublicofcascadia", "thisone")
  
  dataframe %>%
    mutate(country = str_remove_all(country, "[0-9]*"),
         country = str_to_lower(country),
         country = str_remove_all(country, "[`.' !?0-9]"),
         country = if_else(str_detect(country, "uniteds"),"USA", country),
         country = if_else(str_detect(country, "usa"),"USA", country)) %>% 
    mutate(country = case_when(
      country %in% us_list ~ "USA",
      country %in% uk_list ~ "UK",
      country %in% canada_list ~ "Canada",
      country %in% exclude_list ~ NA_character_,
      !is.na(country) & country != "" ~ "Other"
      ), .after = country)
}


```

# Bind tables together

```{r}

candy_cleaned_2015 %>% 
  bind_rows(candy_cleaned_2016) %>% 
  bind_rows(candy_cleaned_2017) %>% 
  distinct(candy_name) %>% 
  arrange(candy_name)

```
```{r}
candy_cleaned_2015
candy_cleaned_2016
candy_cleaned_2017
```



Analysis questions
What is the total number of candy ratings given across the three years. (Number of candy ratings, not the number of raters. Don’t count missing values)
#just a list of all the candy ratings
What was the average age of people who are going out trick or treating?
# age column
What was the average age of people who are not going trick or treating?
# age column
For each of joy, despair and meh, which candy bar received the most of these ratings?
# candy ratings
How many people rated Starburst as despair?
# starburst column (candy columns)

For the next three questions, count despair as -1, joy as +1, and meh as 0.

What was the most popular candy bar by this rating system for each gender in the dataset ?
# candy ratings
What was the most popular candy bar in each year?
# date or year
What was the most popular candy bar by this rating for people in US, Canada, UK, and all other countries?
# countries