---
title: "R Notebook"
output: html_notebook
---

```{r}

library(tidyverse)
library(readxl)
library(janitor)
library(here)
library(assertr)

```

```{r}

candy_2015 <- read_xlsx("raw_data/boing-boing-candy-2015.xlsx") %>%
  clean_names()
candy_2016 <- read_xlsx("raw_data/boing-boing-candy-2016.xlsx") %>%
  clean_names()
candy_2017 <- read_xlsx("raw_data/boing-boing-candy-2017.xlsx") %>%
  clean_names()

```

```{r}

candy_2015
candy_2016
candy_2017


```




```{r}

names_2015 <- names(candy_2015)
names_2016 <- names(candy_2016)
names_2017 <- names(candy_2017)

```

### Clean 2015 data

```{r}

candy_cleaned_2015 <- candy_2015 %>% 
  rename(age = how_old_are_you, 
         going_out = are_you_going_actually_going_trick_or_treating_yourself) %>% 
  mutate(year = str_extract(timestamp, '[0-9]{1,4}'), .after = timestamp) %>%
  mutate(id = row_number(timestamp) + 1e6) %>% 
  mutate(gender = NA_character_, .after = age) %>% 
  mutate(country = NA_character_, .after = age) %>% 
  select(id, year:york_peppermint_patties, necco_wafers) %>% 
  # replace all non integer age inputs as NA, convert values to integers
  mutate(age = as.integer(age), year = as.integer(year)) %>% 
  pivot_longer(butterfinger:necco_wafers, names_to = "candy_name", values_to = "rating") %>%
  clean_candy_names() %>% 
  select(id, year, going_out, age, gender, country, candy_name, rating)

candy_cleaned_2015

```

### Clean 2016 data

```{r}

candy_cleaned_2016 <- candy_2016 %>% 
  rename(going_out = are_you_going_actually_going_trick_or_treating_yourself,
         age = how_old_are_you,
         country = which_country_do_you_live_in,
         gender = your_gender) %>%
  mutate(id = row_number(timestamp) + 2e6, .before = timestamp) %>% 
  mutate(year = str_extract(timestamp, '[0-9]{1,4}'), .after = timestamp) %>%
  clean_country_names() %>% 
  select(id, year:york_peppermint_patties, gender, -which_state_province_county_do_you_live_in) %>% 
  # replace all non integer age inputs as NA, convert values to integers
  mutate(age = as.integer(age), year = as.integer(year)) %>%
  pivot_longer(x100_grand_bar:york_peppermint_patties, names_to = "candy_name", values_to = "rating") %>%
  clean_candy_names() %>% 
  select(id, year, going_out, age, gender, country, candy_name, rating)


candy_cleaned_2016

```

```{r}

candy_cleaned_2017 <- candy_2017 %>% 
  rename(id = internal_id) %>% 
  pivot_longer(q1_going_out:q11_day, names_to = "col_names", values_to = "value") %>%
  select(id, col_names, value) %>%
  mutate(col_names = str_remove(col_names, "q[0-9]_")) %>%
  pivot_wider(names_from = col_names, values_from = value) %>%
  clean_names() %>%
  mutate(year = as.integer(2017), .after = id) %>%
  mutate(age = as.integer(age)) %>%
  clean_country_names() %>% 
  pivot_longer(x100_grand_bar:york_peppermint_patties, names_to = "candy_name", values_to = "rating") %>% 
  clean_candy_names () %>% 
  select(id, year, going_out, age, gender, country, candy_name, rating) %>% 
  filter(!is.na(rating))

candy_cleaned_2017

```

# Bind tables together

```{r}

combined_data_long <- candy_cleaned_2015 %>% 
  bind_rows(candy_cleaned_2016) %>% 
  bind_rows(candy_cleaned_2017) %>% 
  #clean_candy_names() %>% 
  mutate(age = if_else((age > 0 & age < 100), age, NA_integer_)) %>% 
  filter(!is.na(rating))

combined_data_long

combined_data_wide <- combined_data_long %>%
  pivot_wider(names_from = candy_name, values_from = rating)
combined_data_wide

```


--------------------------------------------------------------------------------

Analysis questions
What is the total number of candy ratings given across the three years. (Number of candy ratings, not the number of raters. Don’t count missing values)

```{r}

combined_data_long %>%
  filter(!is.na(rating)) %>% 
  summarise(total_candy_ratings = n())

```

What was the average age of people who are going out trick or treating?
# age column

```{r}

combined_data_wide %>% 
  filter(going_out == "Yes") %>% 
  summarise(avg_age = mean(age, na.rm = TRUE))

```


What was the average age of people who are not going trick or treating?
# age column

```{r}

combined_data_wide %>% 
  filter(going_out == "No") %>% 
  summarise(avg_age = mean(age, na.rm = TRUE))


```


For each of joy, despair and meh, which candy bar received the most of these ratings?
# candy ratings

```{r}

combined_data_long %>%
  group_by(rating, candy_name) %>% 
  summarise(count = n()) %>% 
  slice_max(count) 

```



How many people rated Starburst as despair?
# starburst column (candy columns)

```{r}

combined_data_long %>% 
  filter(candy_name == "starburst", rating == "DESPAIR") %>% 
  summarise(count = n())
  
  
```


For the next three questions, count despair as -1, joy as +1, and meh as 0.



What was the most popular candy bar by this rating system for each gender in the dataset ?
# candy ratings, gender

```{r}

candy_scored <- combined_data_long %>% 
  mutate(score = case_when(
    rating == "JOY" ~ 1,
    rating == "DESPAIR" ~ -1,
    TRUE ~ 0
  ))

```

```{r}

candy_scored %>%
  filter(!is.na(gender)) %>% 
  group_by(gender, candy_name) %>%
  summarise(total_score = sum(score)) %>% 
  slice_max(total_score)
  

```


What was the most popular candy bar in each year?
# date or year

```{r}

candy_scored %>% 
  group_by(year, candy_name) %>% 
  summarise(total_score = sum(score)) %>% 
  slice_max(total_score)


```


What was the most popular candy bar by this rating for people in US, Canada, UK,
and all other countries?

```{r}

candy_scored %>%
  filter(country == c("USA", "Canada", "UK", "Other")) %>% 
  group_by(country, candy_name) %>% 
  summarise(total_score = sum(score)) %>% 
  slice_max(total_score)

```

