---
title: "R Notebook"
output: html_notebook
---

```{r}
# Load in libraries
library(tidyverse)
library(here)

```



```{r}
# Read in data
decathlon <- read_csv(here("clean_data/clean_data.csv"))

decathlon

```

# Who had the longest long jump seen in the data?

```{r}

decathlon %>%
  select(athlete_name, long_jump) %>% 
  filter(long_jump == max(long_jump))

```
The longest long jump distance in the data set was 7.96m by the athlete named
Clay.

# What was the average 100m time in each competition?

```{r}

decathlon %>%
  group_by(competition) %>% 
  summarise(avg_100m = round(mean(x100m), 2))

```
The average 100m time was 11.18 seconds in the Decastar competition and 10.92 
seconds for the OlympicG competition.

# Who had the highest total points across both competitions?

```{r}

decathlon %>% 
  group_by(athlete_name) %>% 
  summarise(total_points = sum(points)) %>% 
  slice_max(total_points, with_ties = FALSE)

```
Whilst not all athletes competed in both competitions the athlete with the 
highest points across both competitions was Sebrle, with 17,110 points.


# What was the shot-put scores for the top three competitors in each 
# competition?

```{r}

decathlon %>% 
  filter(rank <= 3) %>% 
  select(athlete_name, shot_put, rank, competition) %>% 
  arrange(competition, rank)

```

The top three shotput scores for both competitions are shown above. In both 
competitions Sebrle, Clay and Karpov each bagged the top 3 spots in that order,
however each athlete achieved a higher score in the OlympicG competition 
compared to the Decastar competition.

# What was the average points for competitors who ran the 400m in less than 50
# seconds vs. those than ran 400m in more than 50 seconds?

```{r}
decathlon %>% 
  # create new column to categorise times above and below 50s
  mutate(x400m_time_class = if_else(x400m < 50, "Less than 50s",
                                    "More than 50s")) %>% 
  # group by new column
  group_by(x400m_time_class) %>%
  # calculate mean for each group of times
  summarise(avg_points_400m = round(mean(points), 0)) %>% 
  # sort order of results
  arrange(desc(avg_points_400m))

```
The average points per athlete for competitors who ran the 400m in less than 50
seconds was 8,120 whereas athletes who ran the 400m in greater than 50 seconds
averaged a total points of 7,727.
