---
title: "R Notebook"
output: html_notebook
---

```{r}
# load in libraries
library(tidyverse)
library(here)

```



```{r}
# read in cleaned dataset
cake <- read_csv(here("clean_data/clean_data.csv"))
cake

```


# Which cake has the most cocoa in it?

```{r}

cake %>% 
  filter(ingredient == "Cocoa") %>% 
  slice_max(value)

```
The cake with the most cocoa in it is the "One Bowl Chocolate" cake with 10 
tablespoons of cocoa.

# For sponge cake, how many cups of ingredients are used in total?

```{r}

cake %>% 
  group_by(cake) %>% 
  filter(cake == "Sponge", measure == "cup") %>% 
  summarise(num_cups = sum(value))

```
The sponge cake uses a total total of 3.5 cups of ingredients.


# How many ingredients are measured in teaspoons?

**It has been assumed that the question is asking for the number of distinct 
ingredients used within the cake ingredients dataset that are measured using
teaspoons**

```{r}

cake %>% 
  distinct(ingredient, measure) %>%
  filter(measure == "teaspoon") %>% 
  summarise(num_teaspoon_ingredients = n())

```
The number of distinct ingredients within the cake list which use teaspoon as 
the unit of measurement is 8.

# Which cake has the most unique ingredients?

**For the purpose of this question, a 'unique' ingredient has been defined as an 
ingredient which is only used once across all of the cake ingredients. The 'most
unique' cake has been defined as the cake which has the highest number of
'unique' ingredients.**

```{r}

# identify the ingredients which are only used once and save these in a list
unique_ingredients <- cake %>% 
  group_by(ingredient) %>% 
  summarise(count_of_ingredient = n()) %>% 
  filter(count_of_ingredient == 1) %>% 
  select(ingredient) %>% 
  pull()

# determine unique score based on number of ingredients matching with the
# unique list
cake %>% 
  mutate(has_unique_ingredient= ifelse(ingredient %in% unique_ingredients,
                                       T, F)) %>% 
  group_by(cake) %>%
  summarise(uniqueness_score = sum(has_unique_ingredient)) %>% 
  slice_max(uniqueness_score)

```
The most unique cake is the Angel cake which has 2 unique ingredients in it.

# Which ingredients are used only once?

```{r}

cake %>% 
  group_by(ingredient) %>% 
  summarise(count_of_ingredient = n()) %>% 
  filter(count_of_ingredient == 1)

```
The list above shows that there are 8 ingredients which are only used once
across the entire cake ingredient data set.
