---
title: "R Notebook"
output: html_notebook
---

```{r}

library(tidyverse)
library(here)

```

```{r}

bird <- read_csv(here("clean_data/clean_data.csv"))

```

For each question where we ask for bird names, give the bird’s common name, 
scientific name and species abbreviation.

# Which bird had the most individual sightings?

```{r}

bird %>% 
  group_by(common_name, scientific_name, abbreviation) %>%
  summarise(individual_sightings = n()) %>% 
  arrange(desc(individual_sightings))

```
The wandering albatross had the highest number of sightings - 11,247 in total.

# Which bird had the highest total count?

```{r}

bird %>% 
  group_by(common_name, scientific_name, abbreviation) %>% 
  summarise(total_count = sum(count)) %>%
  arrange(desc(total_count)) %>% 
  head(1)

```
The bird with the total highest count was the short tailed shearwater with
982,553 of them spotted.

# Which bird had the highest total count above a latitude of -30?

```{r}

bird %>% 
  filter(lat > -30) %>% 
  group_by(common_name, scientific_name, abbreviation) %>% 
  summarise(total_count = sum(count)) %>%
  arrange(desc(total_count)) %>% 
  head(1)

```
The bird with the total highest count above a latitude of -30 was the wedge 
tailed shearwater with 855 of them spotted.

# How many different types of birds were only ever seen in groups of 1?

```{r}

bird %>% 
  group_by(common_name, scientific_name, abbreviation) %>%
  # determine maximum group size recorded
  summarise(max_seen_at_once = max(count)) %>% 
  # filter on max group size of 1
  filter(max_seen_at_once == 1) %>%
  # count number of bird types only seen in groups of 1
  ungroup() %>% 
  summarise(num_seen_only_groups_of_1 = n())

```
There are 23 different bird types that have only been spotted in groups of 1.

# How many penguins were seen? (Hint: there are many types of penguin)

```{r}
bird %>% 
  mutate(penguin_match = str_detect(common_name, "(?i)Penguin")) %>% 
  filter(penguin_match == TRUE) %>%
  summarise(total_observed = sum(count))

```
There were a total of 158 penguins spotted which consisted of a range of 
penguin types.